%% Example CV - Faried Abu Zaid
%% Comprehensive overview of all competencies, experience, and achievements.
%% Use as a master reference when tailoring role-specific CVs.
%%
%% Compile with: lualatex main_example.tex
%% (pdflatex often fails on modern MiKTeX with fontawesome5 font-expansion errors.)

\documentclass[11pt,a4paper,sans]{moderncv}
\moderncvstyle{banking}
\moderncvcolor{blue}

% Force both first and last name AND section headings to render in moderncv
% blue (color1). Default banking on lualatex+MiKTeX leaves these black, which
% looks inconsistent with the rest of the blue accent scheme.
\renewcommand*{\firstnamestyle}[1]{{\fontsize{34}{36}\bfseries\upshape\color{color1}#1}}
\renewcommand*{\lastnamestyle}[1]{{\fontsize{34}{36}\bfseries\upshape\color{color1}#1}}
\renewcommand*{\sectionstyle}[1]{{\sectionfont\color{color1}#1}}

\usepackage[utf8]{inputenc}
\usepackage{needspace}
\usepackage{academicons}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=blue,
    pdftitle={Faried Abu Zaid - CV},
    pdfpagemode=FullScreen,
}
\usepackage[scale=0.80]{geometry}
\usepackage{import}

% personal data
\name{Faried}{Abu Zaid}
\address{Daphnestr. 32, 81925 Munich, Germany}{}{}
\phone[mobile]{+49 1704902149}
\email{fariedaz@gmail.com}
\extrainfo{\faGlobe~\href{https://fariedabuzaid.github.io}{Website} \quad \faGithub~\href{https://github.com/fariedabuzaid}{GitHub} \quad \aidblp~\href{https://dblp.org/pid/31/10950.html}{DBLP}}

\begin{document}

\makecvtitle

% ============================================================
%     PROFILE STATEMENT
% ============================================================

\vspace{6pt}
\small{AI researcher with a PhD in Computer Science and 10+ years across machine learning research, technical leadership, and applied AI. Core strengths in generative modeling (flow matching, normalizing flows), scientific machine learning, Bayesian methods, uncertainty quantification, and trustworthy AI. First-authored VeriFlow (AAAI 2026) and On Uniformly Scaling Flows; contributor to the open-source sbi reloaded toolkit. I ground systems in sound methodology and push them toward real-world impact, most recently at DeepL on real-time speech-to-speech translation at 10B+ parameter scale.}

% ============================================================
%     CORE COMPETENCIES
% ============================================================

\section{Core Competencies}
\vspace{1pt}
\begin{itemize}

\item \textbf{Scientific ML}: simulation-based inference, Bayesian methods, uncertainty quantification, anomaly detection, inverse problems

\item \textbf{Generative modeling}: flow matching, normalizing flows, diffusion models, ODE-based flow modeling

\item \textbf{Trustworthy / responsible AI}: explainable AI (XAI), neural network verification, model validation, failure-mode analysis

\item \textbf{Programming \& ML engineering}: Python (6+ yrs), PyTorch, JAX, scikit-learn, Pyro, sbi, NumPy/SciPy/Pandas, Ray, WandB; reproducible workflows, ablation studies, benchmark design

\item \textbf{Infrastructure}: distributed training/inference, Docker, OpenShift, AWS, GCP, Git

\end{itemize}

% ============================================================
%     PROFESSIONAL EXPERIENCE
% ============================================================

\section{Professional Experience}
\vspace{3pt}
\begin{itemize}

\needspace{5\baselineskip}
\item{\cventry{Jan--Jun 2026}{Researcher (Real-Time Speech-to-Speech Translation)}{DeepL}{Germany}{}{\vspace{1pt}
\begin{itemize}
    \item Contribute to large-scale production ML for DeepL's Voice products in highly distributed environments (training, evaluation, robust deployment)
    \item Work on transformer-based encoder-decoder models with experiment-driven iteration for quality-critical, user-facing applications
    \item Distributed training and optimization of 10B+ parameter models under strict quality, latency, and robustness constraints
\end{itemize}}}

\vspace{3pt}

\needspace{5\baselineskip}
\item{\cventry{2020--2025}{Senior AI Researcher, TransferLab}{appliedAI / appliedAI Institute for Europe}{Munich}{}{\vspace{1pt}
\begin{itemize}
    \item Led a subgroup on uncertainty quantification within the TransferLab, appliedAI's research department
    \item Implemented flow matching for simulation-based inference; contributed to \emph{sbi reloaded}, an open-source SBI toolkit
    \item Co-supervised a PhD on SBI for material design (joint with BMW and the University of the Bundeswehr)
    \item First-authored \emph{VeriFlow} (AAAI 2026) and \emph{On Uniformly Scaling Flows} (NeurIPS 2026 submission)
    \item Designed and taught practitioner ML courses; scaled flagship courses online (+300\% global reach)
    \item Industrial delivery: UK pharma supply-chain forecasting; hybrid-engine optimization with 7 developers (+15\% simulated fuel efficiency); XAI-guided data frameworks for extrusion manufacturing
\end{itemize}}}

\vspace{3pt}

\needspace{5\baselineskip}
\item{\cventry{2018--2019}{Lead Data Scientist, AI for Data Management}{Camelot Consultants}{}{}{\vspace{1pt}
\begin{itemize}
    \item Led end-to-end analytics and ML delivery for ambiguous industrial problems, from data preparation and modeling to deployment and stakeholder communication
\end{itemize}}}

\vspace{3pt}

\needspace{4\baselineskip}
\item{\cventry{2015--2018}{Postdoctoral Researcher, Automata and Logic}{TU Ilmenau}{}{}{\vspace{1pt}
\begin{itemize}
    \item Research in algorithmic model theory and logic; university lecturing in automata theory, logic, and complexity
\end{itemize}}}

\vspace{3pt}

\needspace{4\baselineskip}
\item{\cventry{2011--2015}{Research Assistant, Mathematical Foundations of CS}{RWTH Aachen}{}{}{\vspace{1pt}
\begin{itemize}
    \item Research and teaching in a DFG-funded project on automatic structures
\end{itemize}}}

\end{itemize}

% ============================================================
%     EDUCATION
% ============================================================

\section{Education}
\vspace{1pt}
\begin{itemize}

\needspace{4\baselineskip}
\item{\cventry{2011--2016}{Dr. rer. nat. (PhD) in Computer Science}{RWTH Aachen}{}{\textit{magna cum laude}}{\vspace{1pt}
Dissertation: ``Algorithmic Solutions via Model Theoretic Interpretations'' (advisors: Grädel, Kuske).
}}

\vspace{3pt}

\item{\cventry{2003--2010}{Diploma in Computer Science}{RWTH Aachen}{}{\textit{with distinction}}{\vspace{1pt}
Thesis: ``Definability in $\omega$-automatic structures.'' Focus: logic and complexity theory; minor in Psychology.
}}

\end{itemize}

% ============================================================
%     LANGUAGES
% ============================================================

\section{Languages}
\vspace{1pt}
\begin{itemize}
\item German (native), English (fluent), Arabic (A1, learning).
\end{itemize}

% ============================================================
%     SELECTED PUBLICATIONS
% ============================================================

\section{Selected Publications}
\vspace{1pt}
\begin{itemize}
\item \textbf{F. Abu Zaid}, D. Neider, M. Yalçiner (2026). VeriFlow: Modeling Distributions for Neural Network Verification. \emph{AAAI 2026}.
\item J. Boelts, M. Deistler, \ldots \textbf{F. Abu Zaid}, et al. (2025). sbi reloaded: a toolkit for simulation-based inference workflows. \emph{J. Open Source Software}.
\item \textbf{F. Abu Zaid}, T. Katzke, E. Müller, D. Neider (2025). On Uniformly Scaling Flows. \emph{CoRR abs/2510.09452} (NeurIPS 2026 submission).
\item \textbf{F. Abu Zaid}, E. Grädel, L. Kaiser, W. Pakusa (2014). Model-Theoretic Properties of $\omega$-Automatic Structures. \emph{Theory Comput. Syst.} 55(4).
\end{itemize}

% ============================================================
%     HONORS AND AWARDS
% ============================================================

\section{Grants, Honors \& Professional Memberships}
\vspace{1pt}
\begin{itemize}
\item{Bavarian research grant -- Bavarian AI Act Accelerator (2024--2025).}
\item{DFG research grant -- Automatic Structures (2012--2015).}
\item{PhD in Computer Science, \textit{magna cum laude} (2016).}
\item{Diploma in Computer Science, with distinction (2010).}
\item{Member of the Association for Computing Machinery (ACM) since 2020.}
\end{itemize}

% Note: no "References available upon request" filler section - the header
% links DBLP/GitHub and references are provided only if actually listed.

\end{document}